Microlensing as a possible probe of event-horizon structure in quasars

### 2017-09-11 paper1/ms.tex

In quasars which are lensed by galaxies, the point-like images sometimes show sharp and uncorrelated brightness 
variations (microlensing). These brightness changes are associated with the innermost region of the quasar passing 
through a complicated pattern of caustics produced by the stars in the lensing galaxy. In this paper, we study 
whether the universal properties of optical caustics could enable extraction of shape information about the central 
engine of quasars. We present a toy model with a crescent-shaped source crossing a fold caustic. The silhouette 
of a black hole over an accretion disk tends to produce roughly crescent sources. When a crescent-shaped source 
crosses a fold caustic, the resulting light curve is noticeably different from the case of a circular luminosity 
profile or Gaussian source. With good enough monitoring data, the crescent parameters, apart from one degeneracy, 
can be recovered.

@@

In Section~\ref{sec:numerics} we carry out source fitting to
lightcurves, with both noise and systematic errors are present.  We
generate a mock lightcurve by taking a crescent source across the path
CB in Figure~\ref{fig:magnification_map}, and then adding noise.  The
path simulates crossing a clean but not ideal fold.  We then fit this
lightcurve to templates from both crescent and circular-Gaussian
sources along the paths AB as well as CB.  That is, the templates from
a similar but not identical caustic are also used, thus deliberately
generating a systematic error.  Despite this systematic error, a
crescent source fits the data, while a Gaussian source is rejected
from the $\chi^2$ value.

@@

\begin{figure}
\includegraphics[width = .49\textwidth]{figures/infinite_fold_ar.eps}
\caption{\label{fig:infinite_fold} Source profiles for Gaussian, disk,
  crescents (right) and magnification map for an infinite fold
  (left). Objects have the same $S_{0}$ and $r_{1/2}$. The black arrow marks the direction of motion of the
sources relative to the caustic.}
\end{figure}

@@

\subsection{Disk source with constant surface brightness distribution}

@@

\begin{equation}
 S_{1D}^D(p-p_s) = \frac{2 S_0^D}{\pi R}  \sqrt{1 - \frac{(p-p_s)^2}{R^2} }    \Theta \left( R^2 - \left( p-p_s \right)^2 \right).
\end{equation}
The half-light radius of a uniform disk source is $R/\sqrt{2}$.

@@

The surface brightness distribution of a geometric crescent can be built 
by considering two disk sources of constant brightness. One larger disk 
will contribute positively to the total flux while one smaller disk 
that is interior to the large one will contribute negatively. This superposition can be written for 2D as:\\

@@

\begin{equation}
 S_{2D}^{Dp}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dp}}{\pi R_p^2} \Theta \left( R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right)
\end{equation}
\\
and
\begin{equation}
 S_{2D}^{Dn}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dn}}{\pi R_n^2} \Theta \left( R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right).
\end{equation}
The following notations were used: $R_p$ and $(p_{sp}, q_{sp})$ are
the radii and coordinate of the center for the larger positive disk, 
while $R_n$ $(p_{sn},q_{sn})$ correspond to the smaller negative disk. 
$S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk. 
From this point forward we will not use the total flux from each source. Instead we will 
use the difference which in this case is the total flux from the crescent-shaped source $S_0^C$. \\
Equation \ref{eqn:s2d} can be written as:\\

@@

There are some constraints on the parameters used to define a crescent
in the previously presented manner that need to be stated. First, we
must impose the obvious $R_p \geq R_n$ relation. Secondly, the small disk
must always be interior to the large disk:
\begin{equation}
 R_p \ge R_n + \sqrt{\left(p_{sp} - p_{sn} \right)^2 + \left(q_{sp} - q_{sn} \right)^2}
\end{equation}
For the distances between the centers of the two disks we will use the same 
notations as the one found in the paper \citep{2013MNRAS.434..765K}, $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.

The half-light radius of any source is invariant under any rotational
transformation.  In the present case of a crescent source the
effective radius is dependent on the parameters $R_p$, $R_n$ and
$\sqrt{a^2+b^2}\equiv c $ exclusively. From symmetry considerations
the centroid of the source is collinear with the centers of the two
disks and it is situated at a distance $d_c$ from the center of the
bright disk. $d_c$ can be computed numerically with the use of a
variation of equation \ref{eqn:s1_d}:

@@

\begin{figure}
\includegraphics[width = .49\textwidth]{figures/figure_rhalf2.eps}
\caption{\label{fig:geom_crescent} Geometry of crescent sources: the source can be seen as a superposition
 of a smaller dark disk over and inside a larger brighter disk. $R_p$ and $R_n$ mark the radii of the 
bright and dark disk. $d_c$ and $c$ mark the distance from the center of the bright disk to the 
centroid and from the dark disk to the centroid.  }
\end{figure}

@@

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/S1D_all.eps}
    \includegraphics[width = 0.48\textwidth]{figures/4source_magnification.eps}
\caption{\label{fig:lightcurve_gauss} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent, Gaussian and disk sources with identical total flux ($S_0$) and half-light radius($r_{1/2}$). Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$. }
\end{figure}

\subsection{Lightcurve of the disk shaped source}

@@

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_a.eps}
    \includegraphics[width = 0.48\textwidth]{figures/4avar_magnification.eps}
\caption{\label{fig:a_var} 1D profiles (top panel) and simulated lightcurves 
(bottom) of crescent sources with identical $R_p$ and $R_n$ ($R_n$ = 0.4 $R_p$) and different disk centers displacement (a).}
\end{figure}

@@

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_poz.eps}
    \includegraphics[width = 0.48\textwidth]{figures/5Rn_back_var_magnification.eps}
\caption{\label{fig:lightcurve_crescent_back} 1D profiles (top panel) and simulated lightcurves (bottom) of backwards crescent sources with identical $R_p$ and $a$ ($a$ = 0.3 $R_p$) and different dark disk radii ($R_n$).}
\end{figure}

@@

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_neg.eps}
    \includegraphics[width = 0.48\textwidth]{figures/5Rn_forw_var_magnification.eps}
\caption{\label{fig:lightcurve_crescent} 1D profiles (top panel) and simulated lightcurves (bottom) 
of forwards crescent sources with identical $R_p$ and $a$ ($a$ = - 0.3 $R_p$) and different dark disk radii ($R_n$). }
\end{figure}

@@

Figure \ref{fig:lightcurve_gauss} reveals that the 
lightcurve of a crescent source has more
visible features than the other two light-curves corresponding to the
disc and Gaussian shape. There are three regions where $S_{1D}$ and by
inheritance, the lightcurve has distinguishable behaviour. The first
region that would be recorded on a lightcurve plot represents the
period of time when the bright disk begins to be overlapped by the
caustic and stops when just before the dark disk reaches the
caustic. During this period of time, the flux of light from the source
is increasingly magnified. The second period starts and ends with the
overlapping of the dark disc. As a boundary of the two regions, there
is a distinguishable point where the slope of the magnification is
drastically changed. This apparent discontinuity in the first
derivative of the magnification function is caused by the caustic
amplification of the sudden drop in the $S_{1d}$ function. During the
respective period, the magnification growth slows or even reverses
and later starts to grow faster again as
the dark disc ends its overlap with the caustic. At the point where
the dark disc clears the caustic, the growth of the magnification is
infinite, which appears as a saddle point on the lightcurve. Next, the
final period corresponds to the case when the dark disc has cleared
the caustic and the bright disc continues to overlap with the
caustic. During this period, the lightcurve reaches a peak that for
most of the parameter space is global and for the rest of the
parameter space local.  At the end of the period, the growth of the
magnification is negative infinite. The respective point appears as a
second saddle point on the lightcurve. Past this point the
magnification of any finite source will decay in roughly the same
manner.

The impact of parameter $a$ on the shape of the lightcurve can be
observed in Figure~\ref{fig:a_var}. Sources where the center of the
dark disc reaches the caustic before the center of the bright disc are
characterized by a smoother broad peak in contrast to the cases where
the center of the dark disc reaches the caustic after the bright disk.
In the case of the latter the instance when the dark disc reaches the
caustic corresponds to larger and larger magnifications until it
becomes a local and even global peak. The effect of the $R_n$
parameter on the light curve is presented in
Figures~\ref{fig:lightcurve_crescent_back} and
\ref{fig:lightcurve_crescent}. For the particular set of
parameters where $R_p = R_n +a$ the third period of time discussed
previously does not exist. Particularizing further, if the value of
the radius of the dark disc is comparable to the value of the radius
of the bright one then the position and shape of the maximum
magnification are strikingly different. In case the crescent reaches
the caustic with the bright region first, the peak magnification
happens when the dark region reaches the caustic and it is
characterized by a sharp variation in magnification growth. In the
opposite case, the peak appears before the end of the bright disc
reaches the caustic and the shape of the lightcurve is smoother.

@@

@@

The mock data appear in Figures
\ref{fig:cc_forward}--\ref{fig:gc_forward}.  The example values
above for $r_{1/2}$, $v_p$ and the observing frequency are not
important.  What is important is that the observations have enough
sampling and signal-to-noise that the characteristic features of a
crescent source (see e.g., Figure~\ref{fig:char_points}) are not
washed out.

@@

The first two source profiles that we consider are the uniform disk and the symmetric Gaussian source. 
Both of them can be described by a half-light radius ($r_{1/2}$) and a total unlensed light flux ($S_0$). 
With the two parameters constrained the one-dimensional profiles, as well as the lightcurves of the 
two source, are completely determined, since no free parameter remains.

The previous statement does not hold for a crescent source.
In the case of the crescent source there are in total five parameters: the integrated flux of the source $S_0$,
 the radii of the bright and dark disks and the displacement of the centers of the two disks on the axes perpendicular and parallel to the caustic.

@@

Since the one-dimensional source profile that corresponds to an 
axis perpendicular to the caustic contains exhaustively all the 
information regarding the source that can be revealed by the lightcurve, 
the value of the crescent's disks centers displacement parallel to the caustic 
does not have an effect on the shape of the lightcurve.

Nevertheless, the respective displacement is relevant for the calculation of
of the half-light radius.  Its qualitative effect is to 
decrease the value of the
half-light radius when the absolute value of the parameter is
increased.  In contrast to the simpler source shapes discussed before the
crescent has two additional relevant parameters: the radius of the dark 
disk ($R_n$) and the center displacement in the direction
 perpendicular to the caustic $a$.

Figure \ref{fig:lightcurve_gauss} reveals that the lightcurve of a crescent source has more
visible features than the other two light-curves corresponding to the
disc and Gaussian shape. The parameters $R_n$ and $a$ have strong
influences on the shape of the microlensed lightcurve as can be seen
in Figures~\ref{fig:a_var}, \ref{fig:lightcurve_crescent_back} and
\ref{fig:lightcurve_crescent}.  Moreover, the one-dimensional
source profile corresponding to the direction perpendicular to the
caustic reveal four characteristic points. The overlap of each of
these points with the caustic leaves visible features on the
lightcurve at the corresponding instances of time. In time order,
the instances correspond to the start of the overlap between the
caustic and the bright disk, the start of the overlap between the
caustic and the dark disk, the end of the overlap between the caustic
and dark disk and finally the end of the overlap between the caustic
and the bright disk.

@@

The previously mentioned abstract parameters can be related to
physical quantities specific to the central region of a quasar.  As
such the luminous region would correspond to the bright accretion disc
that surrounds the black hole. The later's gravity would cause a
shadow in the bright region limited by the extent of the event horizon
of the black hole. Therefore, the radius of the bright disk would
provide an estimate of the size of the accretion disk and the radius
of the dark disc would provide and estimate of the gravitationally
magnified Schwarzschild radius of the black hole $R_{S}^{magnified} =
\Delta t_{dark} \cdot v_p$.  By $R_{S}^{magnified}$ we refer to the
apparent Schwarzschild radius which is larger than the real value at
large distances due to the black hole's own gravity.  Moreover, the
gravitationally magnified value of the Schwarzschild radius is a
monotonic function of the black hole's mass. Therefore, it can be used
to estimate the mass of the black hole if it was not rotating. In the
previous expression, the $\Delta t_{dark}$ denotes the period of
overlap between the black hole shadow and fold caustic, while $v_p$
denotes the component of the relative velocity of the source and fold
which is perpendicular to the caustic.  The respective velocity is an
unknown, though it can be constrained on a case by case basis to an
order of magnitude or even better. This would require the study of the
dynamics of the stellar structure which contains the gravitational
lens and the bulk transverse velocity of the galaxy with respect to
the observer.  A better estimate of the relative velocity would
facilitate a better estimate of the effective non-rotating black hole
mass associated with the black hole shadow.

>>>>

In quasars which are lensed by galaxies, the point-like images sometimes show sharp and uncorrelated brightness 
variations (microlensing). These brightness changes are associated with the innermost region of the quasar passing 
through a complicated pattern of caustics produced by the stars in the lensing galaxy. In this paper, we study 
whether the universal properties of optical caustics could enable extraction of shape information about the central 
engine of quasars. We present a toy model with a crescent-shaped source crossing a fold caustic. The silhouette 
of a black hole over an accretion disc tends to produce roughly crescent sources. When a crescent-shaped source 
crosses a fold caustic, the resulting light curve is noticeably different from the case of a circular luminosity 
profile or Gaussian source. With good enough monitoring data, the crescent parameters, apart from one degeneracy, 
can be recovered.

@@

In Section~\ref{sec:numerics} we carry out source fitting to
lightcurves, where both noise and systematic errors are present.  We
generate a mock lightcurve by taking a crescent source across the path
CB in Figure~\ref{fig:magnification_map}, and then adding noise.  The
path simulates crossing a clean but not ideal fold.  We then fit this
lightcurve to templates from both crescent and circular-Gaussian
sources along the paths AB as well as CB.  That is, the templates from
a similar but not identical caustic are also used, thus deliberately
generating a systematic error.  Despite this systematic error, a
crescent source fits the data, while a Gaussian source is rejected
from the $\chi^2$ value.

@@

\begin{figure}
\includegraphics[width = .49\textwidth]{figures/infinite_fold_ar.eps}
\caption{\label{fig:infinite_fold} Source profiles for Gaussian, disc,
  crescents (right) and magnification map for an infinite fold
  (left). Objects have the same $S_{0}$ and $r_{1/2}$. The black arrow marks the direction of motion of the
sources relative to the caustic.}
\end{figure}

@@

\subsection{Disc source with constant surface brightness distribution}

@@

\begin{equation}
 S_{1D}^D(p-p_s) = \frac{2 S_0^D}{\pi R}  \sqrt{1 - \frac{(p-p_s)^2}{R^2} }    \Theta \left( R^2 - \left( p-p_s \right)^2 \right).
\end{equation}
The half-light radius of a uniform disc source is $R/\sqrt{2}$.

@@

The surface brightness distribution of a geometric crescent can be built 
by considering two disc sources of constant brightness. One larger disc 
will contribute positively to the total flux while one smaller disc 
that is interior to the large one will contribute negatively. This superposition can be written for 2D as:\\

@@

\begin{equation}
 S_{2D}^{Dp}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dp}}{\pi R_p^2} \Theta \left( R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right)
\end{equation}
\\
and
\begin{equation}
 S_{2D}^{Dn}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dn}}{\pi R_n^2} \Theta \left( R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right).
\end{equation}
The following notations were used: $R_p$ and $(p_{sp}, q_{sp})$ are
the radii and coordinate of the center for the larger positive disc, 
while $R_n$ $(p_{sn},q_{sn})$ correspond to the smaller negative disc. 
$S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disc. 
From this point forward we will not use the total flux from each source. Instead we will 
use the difference which in this case is the total flux from the crescent-shaped source $S_0^C$. \\
Equation \ref{eqn:s2d} can be written as:\\

@@

There are some constraints on the parameters used to define a crescent
in the previously presented manner that need to be stated. First, we
must impose the obvious $R_p \geq R_n$ relation. Secondly, the small disc
must always be interior to the large disc:
\begin{equation}
 R_p \ge R_n + \sqrt{\left(p_{sp} - p_{sn} \right)^2 + \left(q_{sp} - q_{sn} \right)^2}
\end{equation}
For the distances between the centers of the two discs we will use the same 
notations as the one found in \cite{2013MNRAS.434..765K}, $a \equiv p_{sn} - p_{sp}$ and $b \equiv q_{sn} - q_{sp}$.

The half-light radius of any source is invariant under any rotational
transformation.  In the present case of a crescent source the
effective radius is dependent on the parameters $R_p$, $R_n$ and
$\sqrt{a^2+b^2}\equiv c $ exclusively. From symmetry considerations
the centroid of the source is collinear with the centers of the two
discs and it is situated at a distance $d_c$ from the center of the
bright disc. $d_c$ can be computed numerically with the use of a
variation of equation \ref{eqn:s1_d}:

@@

\begin{figure}
\includegraphics[width = .49\textwidth]{figures/figure_rhalf2.eps}
\caption{\label{fig:geom_crescent} Geometry of crescent sources: the source can be seen as a superposition
 of a smaller dark disc over and inside a larger brighter disc. $R_p$ and $R_n$ mark the radii of the 
bright and dark disc. $d_c$ and $c$ mark the distance from the center of the bright disc to the 
centroid and from the dark disc to the centroid.  }
\end{figure}

@@

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/S1D_all.eps}
    \includegraphics[width = 0.48\textwidth]{figures/4source_magnification.eps}
\caption{\label{fig:lightcurve_gauss} 1D profiles (top panel) and simulated lightcurves (bottom) of crescent, Gaussian and disc sources with identical total flux ($S_0$) and half-light radius($r_{1/2}$). Crescent source has $r_{1/2}$ = 0.72754 $R_p$, $R_n$ = 0.4 $R_p$, a = 0.3/-0.3 $R_p$. }
\end{figure}

\subsection{Lightcurve of the disc shaped source}

@@

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_a.eps}
    \includegraphics[width = 0.48\textwidth]{figures/4avar_magnification.eps}
\caption{\label{fig:a_var} 1D profiles (top panel) and simulated lightcurves 
(bottom) of crescent sources with identical $R_p$ and $R_n$ ($R_n$ = 0.4 $R_p$) and different disc centers displacement (a).}
\end{figure}

@@

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_poz.eps}
    \includegraphics[width = 0.48\textwidth]{figures/5Rn_back_var_magnification.eps}
\caption{\label{fig:lightcurve_crescent_back} 1D profiles (top panel) and simulated lightcurves (bottom) of backwards crescent sources with identical $R_p$ and $a$ ($a$ = 0.3 $R_p$) and different dark disc radii ($R_n$).}
\end{figure}

@@

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/S1D_var_rn_a_neg.eps}
    \includegraphics[width = 0.48\textwidth]{figures/5Rn_forw_var_magnification.eps}
\caption{\label{fig:lightcurve_crescent} 1D profiles (top panel) and simulated lightcurves (bottom) 
of forwards crescent sources with identical $R_p$ and $a$ ($a$ = - 0.3 $R_p$) and different dark disc radii ($R_n$). }
\end{figure}

@@

Figure \ref{fig:lightcurve_gauss} reveals that the 
lightcurve of a crescent source has more
visible features than the other two light-curves corresponding to the
disc and Gaussian shape. There are three regions where $S_{1D}$ and by
inheritance, the lightcurve has distinguishable behaviour. The first
region that would be recorded on a lightcurve plot represents the
period of time when the bright disc begins to be overlapped by the
caustic and stops when just before the dark disc reaches the
caustic. During this period of time, the flux of light from the source
is increasingly magnified. The second period starts and ends with the
overlapping of the dark disc. As a boundary of the two regions, there
is a distinguishable point where the slope of the magnification is
drastically changed. This apparent discontinuity in the first
derivative of the magnification function is caused by the caustic
amplification of the sudden drop in the $S_{1d}$ function. During the
respective period, the magnification growth slows or even reverses
and later starts to grow faster again as
the dark disc ends its overlap with the caustic. At the point where
the dark disc clears the caustic, the growth of the magnification is
infinite, which appears as a saddle point on the lightcurve. Next, the
final period corresponds to the case when the dark disc has cleared
the caustic and the bright disc continues to overlap with the
caustic. During this period, the lightcurve reaches a peak that for
most of the parameter space is global and for the rest of the
parameter space local.  At the end of the period, the growth of the
magnification is negative infinite. The respective point appears as a
second saddle point on the lightcurve. Past this point the
magnification of any finite source will decay in roughly the same
manner.

The impact of parameter $a$ on the shape of the lightcurve can be
observed in Figure~\ref{fig:a_var}. Sources where the center of the
dark disc reaches the caustic before the center of the bright disc are
characterized by a smoother broad peak in contrast to the cases where
the center of the dark disc reaches the caustic after the bright disc.
In the case of the latter the instance when the dark disc reaches the
caustic corresponds to larger and larger magnifications until it
becomes a local and even global peak. The effect of the $R_n$
parameter on the light curve is presented in
Figures~\ref{fig:lightcurve_crescent_back} and
\ref{fig:lightcurve_crescent}. For the particular set of
parameters where $R_p = R_n +a$ the third period of time discussed
previously does not exist. Particularizing further, if the value of
the radius of the dark disc is comparable to the value of the radius
of the bright one then the position and shape of the maximum
magnification are strikingly different. In case the crescent reaches
the caustic with the bright region first, the peak magnification
happens when the dark region reaches the caustic and it is
characterized by a sharp variation in magnification growth. In the
opposite case, the peak appears before the end of the bright disc
reaches the caustic and the shape of the lightcurve is smoother.

@@

\begin{figure}
\centering
  \includegraphics[width=0.5\textwidth]{compare-mag.eps}
\caption{``Visiting Figure''.  Only in red/blue version, not for the paper.}
\end{figure}

@@

The mock data appear in Figures
\ref{fig:cc_forward}--\ref{fig:gc_forward}.  Illustrative values have
been adopted for the frequency of observations and for the noise
level.  It is, of course, important is that the observations have
enough sampling and signal-to-noise that the characteristic features
of a crescent source (see e.g., Figure~\ref{fig:char_points}) are not
washed out.

@@

The first two source profiles that we consider are the uniform disc and the symmetric Gaussian source. 
Both of them can be described by a half-light radius ($r_{1/2}$) and a total unlensed light flux ($S_0$). 
With the two parameters constrained the one-dimensional profiles, as well as the lightcurves of the 
two source, are completely determined, since no free parameter remains.

The previous statement does not hold for a crescent source.
In the case of the crescent source there are in total five parameters: the integrated flux of the source $S_0$,
 the radii of the bright and dark discs and the displacement of the centers of the two discs on the axes perpendicular and parallel to the caustic.

@@

Since the one-dimensional source profile that corresponds to an 
axis perpendicular to the caustic contains exhaustively all the 
information regarding the source that can be revealed by the lightcurve, 
the value of the crescent's discs centers displacement parallel to the caustic 
does not have an effect on the shape of the lightcurve.

Nevertheless, the respective displacement is relevant for the calculation of
of the half-light radius.  Its qualitative effect is to 
decrease the value of the
half-light radius when the absolute value of the parameter is
increased.  In contrast to the simpler source shapes discussed before the
crescent has two additional relevant parameters: the radius of the dark 
disc ($R_n$) and the center displacement in the direction
 perpendicular to the caustic $a$.

Figure \ref{fig:lightcurve_gauss} reveals that the lightcurve of a crescent source has more
visible features than the other two light-curves corresponding to the
disc and Gaussian shape. The parameters $R_n$ and $a$ have strong
influences on the shape of the microlensed lightcurve as can be seen
in Figures~\ref{fig:a_var}, \ref{fig:lightcurve_crescent_back} and
\ref{fig:lightcurve_crescent}.  Moreover, the one-dimensional
source profile corresponding to the direction perpendicular to the
caustic reveal four characteristic points. The overlap of each of
these points with the caustic leaves visible features on the
lightcurve at the corresponding instances of time. In time order,
the instances correspond to the start of the overlap between the
caustic and the bright disc, the start of the overlap between the
caustic and the dark disc, the end of the overlap between the caustic
and dark disc and finally the end of the overlap between the caustic
and the bright disc.

@@

The previously mentioned abstract parameters can be related to
physical quantities specific to the central region of a quasar.  As
such the luminous region would correspond to the bright accretion disc
that surrounds the black hole. The later's gravity would cause a
shadow in the bright region limited by the extent of the event horizon
of the black hole. Therefore, the radius of the bright disc would
provide an estimate of the size of the accretion disc and the radius
of the dark disc would provide and estimate of the gravitationally
magnified Schwarzschild radius of the black hole $R_{S}^{magnified} =
\Delta t_{dark} \cdot v_p$.  By $R_{S}^{magnified}$ we refer to the
apparent Schwarzschild radius which is larger than the real value at
large distances due to the black hole's own gravity.  Moreover, the
gravitationally magnified value of the Schwarzschild radius is a
monotonic function of the black hole's mass. Therefore, it can be used
to estimate the mass of the black hole if it was not rotating. In the
previous expression, the $\Delta t_{dark}$ denotes the period of
overlap between the black hole shadow and fold caustic, while $v_p$
denotes the component of the relative velocity of the source and fold
which is perpendicular to the caustic.  The respective velocity is an
unknown, though it can be constrained on a case by case basis to an
order of magnitude or even better. This would require the study of the
dynamics of the stellar structure which contains the gravitational
lens and the bulk transverse velocity of the galaxy with respect to
the observer.  A better estimate of the relative velocity would
facilitate a better estimate of the effective non-rotating black hole
mass associated with the black hole shadow.
